---

---

<center>  <h3> Fayruz Kibria </h2> </center>

---

#### 1. Check the datatypes of the attributes. (3 points)

```{r}
winedata<-read.csv(file="http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", header=T, sep=";")
#head(winedata)
str(winedata)
```


#### 2. Are there any missing values in the dataset? 

```{r}

#winedata[!complete.cases(winedata),]

cat("There are", sum(is.na(winedata)), "missing data in the dataset.")

```


#### 3. What is the correlation between the attributes other than Quality? 

```{r}
library(tidyr)
library(dplyr)
library(corrplot)

sub.data <-as.data.frame(select( winedata, -quality))

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(cor(sub.data), method="color", col=col(200),  
         type="lower", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=90, number.cex= .6,#Text label color and rotation
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )

```


#### 4. Graph the frequency distribution of wine quality by using Quality. 

```{r message=FALSE, warning=FALSE}
library(ggplot2)
#hist(winedata$quality)

wineC_plot <- ggplot(winedata, aes(quality),fill=quality) +
  geom_histogram(col='black',binwidth=1, boundary = 0.5) +
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10))+
  ylim(0,2500) +
  ggtitle ("Distribution of wine quality") +
  xlab("Wine quality") +
  ylab ("Frequency") +
  theme(plot.title = element_text(hjust = 0.5))

wineC_plot

```


#### 5. Reduce the levels of rating for quality to two levels as Pass and Fail. Assign the levels of 3, 4 and 5 to level Fail; and 6, 7, 8 and 9 to level Pass. 

```{r}
winedata$binaryquality[(winedata$quality <=5)]<- "Fail"
winedata$binaryquality[(winedata$quality >=6)]<- "Pass"
winedata$binaryquality<-as.factor(winedata$binaryquality)

head(winedata$binaryquality)
                   
```


#### 6. Normalize the data set.

```{r}

#standerdizing function function
standardize = function(x){
  z <- (x - mean(x)) / sd(x)
  return( z)
}

#min_max_norm <- function(x) {
#    (x - min(x)) / (max(x) - min(x))
#  }


#apply standerdization to the numeric columns in wine data set and save it as winedata norm dataset
winedata_norm <- as.data.frame(sapply(winedata[1:11], standardize))
#print(max(winedata_norm$residual.sugar))

# add the bimodal class variable to the normalized dataset
winedata_norm$binaryquality <- winedata$binaryquality

#view first six rows of normalized dataset
head(winedata_norm)


```


#### 7. Divide the dataset to training and test sets. 

```{r}

seed = 12345

# creating random sample indices
train_index <- sample(1:nrow(winedata_norm), (2/3) * nrow(winedata_norm))

# seperating train and test data in a 2/3 split
training.data <- winedata_norm[train_index,]
test.data <- winedata_norm[-train_index,]

# random row reordering for test data set
test_index <- sample(1:nrow(test.data), nrow(test.data))
test.data <- test.data[test_index,]

# checking split rows
print(nrow(training.data))
print(nrow(test.data))

```


#### 8. Use the Logistic Regression algorithm to predict the quality of wine using its attributes. 

```{r}

# creating logistic regression model using the train set
logisticRmodel <- glm(binaryquality ~.,family=binomial(link='logit'),data=training.data)
summary(logisticRmodel)

# calculating probabilities using the test set on the logistic regression model
probabilities <-as.data.frame(predict(logisticRmodel,test.data, type = "response"))

# bi modal pass fail level split at 50% probability
predicted.classes <- ifelse(probabilities >= 0.5, "Pass", "Fail")

# storing predicted classes as factors
predicted.classes<- as.factor(predicted.classes)

```


#### 9. Display the confusion matrix to evaluate the model performance. 

```{r}

T_Pass<-sum(predicted.classes[predicted.classes=="Pass"] == test.data$binaryquality[predicted.classes=="Pass"])

F_Fail<-sum(test.data$binaryquality=="Pass")-T_Pass

T_Fail<-sum(predicted.classes[predicted.classes=="Fail"] == test.data$binaryquality[predicted.classes=="Fail"])

F_Pass<-sum(test.data$binaryquality=="Fail")-T_Fail


cat("             Predicted Fail  Predicted Pass")
cat("\n             --------------  --------------")
cat("\nActual Fail |    ",T_Fail, "      |     ",F_Pass,"   |")
cat("\n             --------------  --------------")
cat("\nActual Pass |    ",F_Fail, "      |     ",T_Pass,"   |")
cat("\n             --------------  --------------\n\n")


#confusion_M <- table(actual = test.data$binaryquality, predicted = predicted.classes)
#confusion_M



```


#### 10. Evaluate the model performance by computing Accuracy, Sensitivity and Specificity. 

```{r}

AC <-(T_Pass+T_Fail)/(T_Pass+T_Fail+F_Pass+F_Fail)
SE <-T_Pass/(T_Pass+F_Fail)
SP <-T_Fail/(T_Fail+F_Pass)


cat("Accuracy:    ", round(AC,2))
cat("\nSensitivity: ", round(SE,2))
cat("\nSpecificity: ", round(SP,2))


```